---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Andreas Jeppesen, Thea Pedersen, Sophie Korch & Lina Pedersen"
date: "November 3, 2020" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)


################################# LOADING DATA ####################################
```{r}
library(tidyverse)
library(lme4)
library(caret)


done_df <- read_csv("done_df.csv")

df <- subset(done_df, range_scaled!="NA" & pauseduration_scaled!="NA" & npause!="NA" & phonationtime!="NA" & Diagnosis !="NA")

# m1 <- glmer(Diagnosis ~ range_scaled + (1 | ID), df, family= "binomial")
# 
# summary(m1)
# 
# df$Diagnosis <- ifelse(df$Diagnosis=="1", "Schizophrenia", "Control")
# 
# 
# 
# # We create a function for turning log odds to probabilities:
# logit2prob <- function(logit) {
#   odds <- exp(logit)
#   prob <- odds / (1 + odds)
#   return(prob)
# }
# 
# #The model predicts everyone to be controls. WHY?
# df$pred <- logit2prob(predict(m1, re.form=NA))
# df$pred <- as.numeric(df$pred)
# df$pred <- ifelse(df$pred > 0.5, 1, 0)
# df$pred=as.factor(df$pred)
# df$Diagnosis=as.factor(df$Diagnosis)
# ############################## All comes out control ############################## 
# confusionMatrix(data= df$pred, reference= df$Diagnosis, positive = 1)

```


############################# CREATING FOLDS ######################################
```{r}
#install.packages("groupdata2")
#install.packages("cvms")
library(groupdata2)
library(cvms)

#Ensuring repreducibility
set.seed(1)

df$ID=as.factor(df$ID)

df <- df %>% 
    mutate(
        range_scaled=NULL,
        phonationtime_scaled=NULL,
        speechrate_scaled=NULL,
        pauseduration_scaled=NULL,
        articulation_scaled=NULL)


df <- fold(
  df, 
  k=5, 
  cat_col = "Diagnosis", 
  id_col = "ID", 
  num_fold_cols = 5,
  ) 


# Split into train and test sets
test_df <- df %>% 
  dplyr::filter(.folds_1 == 3)

train_df <- df %>% 
  dplyr::filter(.folds_1 != 3)

 test_df <- na.omit(test_df) 
```



########################### SCALING SPEECH PARAMETERS ##################################
```{r}
test_df <- test_df %>% 
    mutate(
        range_scaled=scale(as.numeric(range)),
        phonationtime_scaled=scale(as.numeric(phonationtime)),
        speechrate_scaled=scale(as.numeric(speechrate)),
        pauseduration_scaled=scale(as.numeric(pauseduration)),
        articulation_scaled=scale(as.numeric(articulationrate))
    )

 train_df <- train_df %>% 
    mutate(
        range_scaled=scale(as.numeric(range)),
        phonationtime_scaled=scale(as.numeric(phonationtime)),
        speechrate_scaled=scale(as.numeric(speechrate)),
        pauseduration_scaled=scale(as.numeric(pauseduration)),
        articulation_scaled=scale(as.numeric(articulationrate))
    )

```



############################### PREDICTING SCORES ########################################
```{r}
# Fitting SVM model
svm_model <- e1071::svm(
    formula = Diagnosis ~ range_scaled + pauseduration_scaled + articulation_scaled, 
    data = train_df,
    kernel = "linear",
    cost = 10,
    type = "C-classification",
    probability = TRUE)


# Predict scores in the test set
predicted_scores <- predict(
  svm_model,
  newdata = test_df,
  allow.new.levels = TRUE,
  probability = TRUE)

test_df$predicted_scores <- as.numeric(predicted_scores)
test_df$Diagnosis <- as.numeric(test_df$Diagnosis)


test_df$Diagnosis=as.factor(test_df$Diagnosis)
test_df$predicted_scores=as.factor(test_df$predicted_scores)

#Creating a confusion matrix
confusionMatrix(data= test_df$predicted_scores, reference= test_df$Diagnosis, positive = "1") 
#Accuracy was 0.59

```



################################ CREATING A MODEL FUNCTION #################################
```{r}

clf_svm_model_fn <- function(train_data, formula, hyperparameters) {

  e1071::svm(
    formula = formula,
    data = train_data,
    type = "C-classification",
    probability = TRUE  # Must enable probability here
  )
}


#Scaling speech values in df  
df <- df %>% 
    mutate(
        range_scaled=scale(as.numeric(range)),
        phonationtime_scaled=scale(as.numeric(phonationtime)),
        speechrate_scaled=scale(as.numeric(speechrate)),
        pauseduration_scaled=scale(as.numeric(pauseduration)),
        articulation_scaled=scale(as.numeric(articulationrate))
    )


#We need to omit NA's from the data in order for the crossvalidation function below to work. 
#To avoid deleting more data than needed, we created a "trimmed" dataframe in which we only include the columns that we will need for the model.
df_trimmed <- df %>% 
  select("ID", "Diagnosis", "mean", ".folds_1", ".folds_2", ".folds_3", ".folds_4", ".folds_5", "pauseduration_scaled", "speechrate_scaled", "range_scaled", "articulation_scaled") %>% 
  na.omit()


#Trying the model function
m1 <- clf_svm_model_fn(train_data = df, formula = Diagnosis ~ range_scaled)
m1

```



############################## CREATING A PREDICT FUNCTION ############################
```{r}
# Predict function for binomial SVM
bnml_svm_predict_fn <- function(test_data, model, formula, hyperparameters, train_data) {
  # Predict test set
  predictions <- predict(
    object = model,
    newdata = test_data,
    allow.new.levels = TRUE,
    probability = TRUE
  )
  
  # Extract probabilities
  probabilities <- dplyr::as_tibble(attr(predictions, "probabilities"))
  
  # Return second column
  probabilities[[2]]
}

#Trying out the predict function
p1 <- bnml_svm_predict_fn(test_data = df, model = m1)
p1    # Vector with probabilities that diagnosis is 1


```



############################# CROSS VALIDATION #################################
```{r}

cv <- cross_validate_fn(
  data = df_trimmed,
  formulas = c("Diagnosis ~ range_scaled + pauseduration_scaled + speechrate_scaled + articulation_scaled + mean",
               "Diagnosis ~ range_scaled + pauseduration_scaled + speechrate_scaled + articulation_scaled",
               "Diagnosis ~ range_scaled + pauseduration_scaled + speechrate_scaled + mean",
               "Diagnosis ~ range_scaled + pauseduration_scaled + articulation_scaled + mean",
               "Diagnosis ~ range_scaled + speechrate_scaled + articulation_scaled + mean", 
              "Diagnosis ~ pauseduration_scaled + speechrate_scaled + articulation_scaled + mean", 
              "Diagnosis ~ range_scaled + articulation_scaled + pauseduration_scaled"),
  type = "binomial",
  model_fn = clf_svm_model_fn,
  predict_fn = bnml_svm_predict_fn, 
  fold_cols = paste0(".folds_", 1:5)
)

#Getting output
cv

#Making it look nicer
cv %>% 
  dplyr::mutate(`Model ID` = 1:nrow(cv_4)) %>% 
  dplyr::arrange(dplyr::desc(`Balanced Accuracy`)) %>% 
  select_definitions(additional_includes = c("Balanced Accuracy", "Model ID")) %>% 
  knitr::kable()
```





######################### DID NOT GONNA USE THIS - TIDYMODELS ######################################
```{r}

# #install.packages("recipes")
# library(recipes)
# 
# df$ID <- as.factor(df$ID)
# 
# df_par <- partition (df, p=0.8, cat_col = "Diagnosis", id_col = "ID", list_out = F)
# df_train <- subset (df_par, .partitions==1)
# df_test <- subset(df_par, .partitions==2)
# 
# rec <- df_train %>% 
#   recipe(Diagnosis ~ ., data= df_train)
# 
# rec_steps <- rec %>% 
#   step_scale(all_numeric()) %>% 
#   step_center(all_numeric())
# 
# prepped_recipe <- prep(rec_steps, training = df_train, retain = T)
# df_train_n <- juice(prepped_recipe)  %>% select(-c(Study,ID,.partitions))
# df_test_n <- bake(rec, new_data = df_test, all_predictors()) %>% select(-c(Study,ID,.partitions))
```

