---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: " Andreas Jeppesen, Thea Pedersen, Sophie Korch & Lina Pedersen"
date: "3/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment (all the other measures have been removed from the dataset for the sake of simplicity). 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
Data: https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0
Prisma chart: https://www.dropbox.com/s/vbjf6ff73dv9tru/PRISMA%202009%20flow%20diagram_PP.pptx?dl=0 
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: PITCH_F0''M and PITCH_F0''SD group of variables are what you need
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

- Now look at the output of rma() and check tau and I2

```{r} 
#Libraries
library(readxl)
library(tidyverse)

#Reading the df
df <- read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")

#Removing all duplications of studies. 
df <- distinct(df, StudyID, .keep_all = TRUE)

# We select only the columns we are interested in: HC= Healthy controls, SZ = Schizophrenia, M=mean, SD = Standart diviation. 
df <- df %>% 
  select(
    ArticleID,
    StudyID,
    Year_publication,
    Article,
    DIAGNOSIS,
    SAMPLE_SIZE_HC,
    SAMPLE_SIZE_SZ,
    PITCH_F0_HC_M,
    PITCH_F0_HC_SD,
    PITCH_F0_SZ_M,
    PITCH_F0_SZ_SD
  )

```

############################# MODELS #####################################

```{r}

# USEFUL LIBRARY
#install.packages("metafor")
library(metafor)


#THIS CREATES TWO NEW COLUMNS yi AND vi. yi= observed effectsize. vi= sampling variance. 
WM_d = escalc(measure = 'SMD', n1i=SAMPLE_SIZE_HC, n2i=SAMPLE_SIZE_SZ, m1i=PITCH_F0_HC_M, m2i=df$PITCH_F0_SZ_M, sd1i=df$PITCH_F0_HC_SD, sd2i=df$PITCH_F0_SZ_SD, data = df)


############################# MODELS: ########################################

# Mixed effects implementation
m <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=WM_d, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m)

# Meta-analysis optimization:
m2 <-rma (yi, vi, data = WM_d, slab = StudyID)
forest(m2)
summary(m2)

# # What if we have a fixed effect?
# m_fixed <- lmer(yi ~ 1 + test+ (1 | study), weights = 1/vi, data=WM_d, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
# m_fixed2 <- rma(yi, vi, mods = cbind(Language), data = data, slab=study)


```

############################# INCLUDING OUR OWN STUDY #####################################

```{r}
# We load the data from A3:
dfA3 <- read_csv("done_df.csv")

# 
dfA3 %>%
   mutate(
        Diagnosis <- as.factor(Diagnosis),
        ID <- as.factor(ID),
        Gender <- as.factor(Gender),
        Study <- as.factor(Study),
    ) %>% 
    split(dfA3$Diagnosis) %>%    # splitting the dataframe into two based on Diagnosis
  map(summary)

# mean pitch controls: 156.8

# sd pitch controls: 
controlsub <- dfA3 %>% subset(Diagnosis == 0)
sd(controlsub$mean)
# 59.19

# mean pitch sz: 161.13

# sd pitch sz:
szsub <- dfA3 %>% subset(Diagnosis == 1)
sd(szsub$mean)
# 56.73


# The number of participants in each diagnosis group
length(unique(controlsub$ID))
# there are 154 controls

length(unique(szsub$ID))


# there are 143 sz patients

# We save our info:  (in this order of variables:)
    #ArticleID,
    # StudyID,
    # Year_publication,
    # Article,
    # DIAGNOSIS,
    # SAMPLE_SIZE_HC,
    # SAMPLE_SIZE_SZ,
    # PITCH_F0_HC_M,
    # PITCH_F0_HC_SD,
    # PITCH_F0_SZ_M,
    # PITCH_F0_SZ_SD,
Studygroup3 <- c(49, 51, 2020, "Studygroup 3 (2020)", "SZ", 154, 143, 156.8, 59.19, 161.13, 56.73)

# we append this list to our df:
df1 <- rbind(df, Studygroup3)
```


############################# MODELS INCLUDING OUR STUDY #####################################

```{r}
library(metafor)

df1 <- df1 %>% 
  mutate(
    SAMPLE_SIZE_HC = as.numeric(SAMPLE_SIZE_HC),
    SAMPLE_SIZE_SZ = as.numeric(SAMPLE_SIZE_SZ),
    PITCH_F0_HC_M = as.numeric(PITCH_F0_HC_M),
    PITCH_F0_SZ_M = as.numeric(PITCH_F0_SZ_M),
    PITCH_F0_HC_SD = as.numeric(PITCH_F0_HC_SD),
    PITCH_F0_SZ_SD = as.numeric(PITCH_F0_SZ_SD)
  )


#THIS CREATES TWO NEW COLUMNS yi AND vi. yi= observed effectsize. vi= sampling variance. 
WM_d1 = escalc(measure = 'SMD', n1i=SAMPLE_SIZE_HC, n2i=SAMPLE_SIZE_SZ, m1i=PITCH_F0_HC_M, m2i=PITCH_F0_SZ_M, sd1i=PITCH_F0_HC_SD, sd2i=PITCH_F0_SZ_SD, data = df1)



############################# MODELS: ########################################

# Mixed effects implementation
 m3 <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=WM_d1, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m3)

# Meta-analysis optimization:
m4 <-rma (yi, vi, data = WM_d1, slab = StudyID)
forest(m4)
summary(m4)

#I^2.  > 75% high variance. 25 – 50 % moderate variance. < 25 % low variance
#τ^2 #the lower the value, the lower the heterogeneity (0 when no heterogeneity)

```


############################# PLOTTING #####################################

```{r}
###################### Forest plots and effect size estimation #########################

#Checking if each study has roughly the same influence
inf <- influence(m4)
print(inf)
plot(inf)


#Plotting effect size against standard error
funnel(m4)

#Statistical measures of checking asymmatry in funnel plot
regtest(m4)
ranktest(m4)


##################################### New functions used in this lecture #####################################
# library (metaphor)
# escalc()
# rma()
# forest()
# funnel()
# influence()
# regtest()
# ranktest()

```

